Prediction of US election with linear model

### 2024-10-21 paper/paper.qmd
made first multi linear model with residuals

library(tidyverse)
library(palmerpenguins)
```

# Introduction {#sec-intro}
Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States. Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result. As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts. However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data. This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.

This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek. As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors. Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics. Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses. We incorporate aggregation by filtering the polls by numeric grade that indicates pollster’s reliability, prediction that incorporates historical data and economic indicators, and hybrid approaches that leverages Bayesian techniques, allowing for a dynamic prediction of the U.S. presidential election.

The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump. The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.

The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]

The remainder of this paper is structured as follows: [update this …]

@@

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
us

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

>>>>

library(tidyverse)
library(modelsummary)
```

# Introduction {#sec-intro}

@@

## Model

# ELECTION DATE: 11 Nov 24
# Day Kamala announced she was running: 21 July 2024
# will start looking at polling data from August

# data manipulation
clean_data <- read_csv("../data/02-analysis_data/cleaned_data.csv") %>% 
  filter(answer %in% c("Harris", "Trump", "Kennedy", "Stein"))
```

```{r}
#| label: tbl-multilinearmodel
#| tbl-cap: "Coefficents for Multilinear Regression Model"
#| echo: false
#| warning: false
#| message: false

multi_lm <- lm(pct ~ transparency_score + pollscore + sample_size + answer + days_taken_from_election, clean_data)

modelsummary(multi_lm)
```

\begin{table}[htbp]
\centering
\caption{Coefficients for Multilinear Regression Model}
\label{tbl-multilinearmodel}
\begin{tabular}{lc}
\hline
Variable & Coefficient \\
\hline
Intercept & 51.371 \\
Transparency Score & -0.218 \\
Poll Score & 0.321 \\
Sample Size & 0.000 \\
Candidate Name & Varies \\
Days from Election & Not shown \\
\hline
\end{tabular}
\end{table}

\begin{equation}
\begin{split}
\text{candidate\_percentage} = & -0.218 \times \text{transparency\_score} + 0.321 \times \text{pollscore} \\
& + 0 \times \text{sample\_size} + \text{candidate\_intercept} \times \text{candidate\_name\_vector} \\
& + 51.371
\end{split}
\end{equation}

From @tbl-multilinearmodel, candiate name is the biggest factor in terms of support. The name was added to the model to give a baseline not to show any correlation. After that, poll score and transparency score are the biggest factors. These scores are two baselines of how reliable and trustable the pollsters are. The sample size had a coefficent of zero, meaning it had no affect on the model's output.

The model also has a very high R^2 value in 0.962 Most of this can be attributed to using the canidate name (anwser) as a variable.

## Residuals

```{r}
#| label: fig-multilinearmodel
#| fig-cap: "Multilinear Regression Model Residuals"
#| echo: false
#| warning: false
#| message: false

ggplot(multi_lm, aes(x = .fitted, y = .resid, color = answer)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  labs(y = "Residual", x = "Predicted Candiate Vote Percentage")

```
Based on @fig-multilinearmodel, there appears to be no patterns in the residuals. This means that a linear model like this could be appropriate. However, there does seem to be a big difference in terms of residual variability based on candidate name.
